---
title: "fastqc_summary"
author: "Matthew Spriggs"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Taking the output from fastq I have a folder called `qc_checks/`. This contains html file and zip files for each of my sequences. I would like to summarize the samples into an easy to read format. This workbook will be used as a playground to design a working function. This function will be added to `utils.R` which is a general R script I will use for generic function for this assignment.

## Set wd
```{r}
base_dir="~/Library/CloudStorage/OneDrive-Illumina,Inc./Documents/Applied Bioinformatics/modules/next_generation_sequencing_informatics/assignment/qc_checks/raw"
setwd(base_dir)
```

```{r}
library(tidyr)
library(dplyr)
```

```{r}
# Find all folders
qc_filepath = paste0(base_dir,'qc_checks')
dir_names = list.dirs(path = qc_filepath, recursive = FALSE)
all_summary = list()
sample_names = c()
for (dir in dir_names){
  print(dir)
  file_path = file.path(dir, 'summary.txt')
  if (file.exists(file_path)){
    data = read.table(file_path,
                      header = FALSE,
                      sep = '\t',
                      col.names = c('flag','metric','sample'))
    # Extract sample name
    sample_name = unique(data$sample)
    sample_name = gsub('_[12].fq.gz', '', sample_name)
    sample_names = append(sample_names, sample_name)
    # Append into list all summary data
    all_summary = append(all_summary, list(data))
  } else {
    warning(paste('File not found:', file_path))
  }
}

names(all_summary) = sample_names
# all_summary is a list with all the summary data
print(all_summary[[1]])
```

```{r}
# Loop through each item in all_summary and pivot

pivoted_list = list()

for (sample_summary in all_summary){
  # print(sample_summary)
  pvt = pivot_wider(sample_summary,names_from = metric, values_from = flag)
  # print(pvt)
  pivoted_list = append(pivoted_list, list(pvt))
  print(length(pivoted_list))
}

# combine into dataframe
df = bind_rows(pivoted_list)
df

```

```{r}
#' Pivot Summary Files
#'
#' This function reads summary files from a specified directory, processes the data, and returns a combined data frame with pivoted data.
#'
#' @param base_dir A string specifying the base directory containing the 'qc_checks' folder.
#' @return A data frame with pivoted summary data.
#' @examples
#' result <- pivot_summary_files("/path/to/your/folder/")

pivot_summary_files = function(base_dir) {
  qc_filepath = base_dir
  dir_names = list.dirs(path = qc_filepath, recursive = FALSE)
  all_summary = list()
  sample_names = c()
  
  for (dir in dir_names) {
    file_path = file.path(dir, 'summary.txt')
    if (file.exists(file_path)) {
      data = read.table(
        file_path, header = FALSE,
        sep = '\t',
        col.names = c('flag', 'metric', 'sample'))
      sample_name = unique(data$sample)
      sample_name = gsub('_[12].fq.gz', '', sample_name)
      sample_names = append(sample_names, sample_name)
      all_summary = append(all_summary, list(data))
    } else {
      warning(paste('File not found:', file_path))
    }
  }
  
  names(all_summary) = sample_names
  pivoted_list = list()
  
  for (sample_summary in all_summary) {
    pvt = pivot_wider(sample_summary, names_from = metric, values_from = flag)
    pivoted_list = append(pivoted_list, list(pvt))
  }
  
  df = bind_rows(pivoted_list)
  return(df)
}

```

```{r}
df = pivot_summary_files(base_dir)
write.table(df, paste0(base_dir,'/raw_summary.csv'), sep = ',', row.names = FALSE)
```